### 2018-06-13 doc/methods.tex
temperature data. just plotting the data for funzies

Microfossil occurrence data was downloaded from the Neptune Database \url{http://www.nsb-mfn-berlin.de/nannotax}. Occurrence information was downloaded for calcareous nannofossils, diatoms, dinoflaggellants, foraminifera, and radiolarians from the entire globe and having occurred after 65 Mya.

@@

@@

\(\Theta\) is then parameterized as a hierarchical model with multiple non-nested varying intercepts CITATION. We considered two aspects of geographic range as continuous covariates: geographic range \(r\) during interval \(t\), and the difference \(d\) between the geographic range at \(t - 1\) and \(t\). We also considered the interaction between geographic range during intervals \(t\) and \(t - 1\). At the first observation for a taxon, the diff of geographic range is 0.

Our model had up to three non-nested varying intercepts: age of observation \(a\), bin of observation \(b\), and taxonomic group of the observation \(c\). Age of observation is a critical varying intercept as this effect, when considered along with the intercept \(\beta_{0}\), captures changes to baseline hazard with species age. Because these factors are modeled as varying intercepts, we are assuming that the effects of geographic range (\(\beta_{1} \dots \beta_{3}\))

>>>>

Microfossil occurrence data was downloaded from the Neptune Database \url{http://www.nsb-mfn-berlin.de/nannotax}. Occurrence information was downloaded for calcareous nannofossils, diatoms, dinoflaggellants, foraminifera, and radiolarians from the entire globe and having occurred between 120 and 0 Mya. This selection of species was then culled to just those species which have their first occurrence at most 63 My; this avoids those taxa which survived the K/Pg boundary, those taxa arrising just after the K/Pg boundary, and lines up the occurrences with the temperature time-series (discussed below).

@@

Temperature data is based on estimates from Cramer et al CITATION. These estimates are based on Magnesian/Calcium isotope ratios and are a more accurate estimate of global temperature than the standard Oxygen isotope-based estimates because Mg/Ca based estimates are not effected by ice-volume and fresh-water input (e.g. metioric water). Cramer et al. provide temperature estimates for every 0.1 My from 0 to 63 Mya; we binned these estimates every 1 My in accordance with the way we binned all fossil occurrences.

@@

\(\Theta\) is then parameterized as a hierarchical model with multiple non-nested varying intercepts CITATION. First, we considered two aspects of geographic range as continuous covariates: geographic range \(r\) during interval \(t\), and the difference \(d\) between the geographic range at \(t - 1\) and \(t\). We also considered the interaction between geographic range during intervals \(t\) and \(t - 1\). At the first observation for a taxon, the diff of geographic range is 0.

Second, we considered two aspects of global temperature: mean temperature during interval \(t\), and the mean temperature during interval \(t - 1\).

Our model had up to three non-nested varying intercepts: age of observation \(a\), bin of observation \(b\), and taxonomic group of the observation \(c\). Age of observation is a critical varying intercept as this effect, when considered along with the intercept \(\beta_{0}\), captures changes to baseline hazard with species age. Because these factors are modeled as varying intercepts, we are assuming that the effects of geographic range (\(\beta_{1} \dots \beta_{3}\)) and temperature (\(\beta_{4} \dots \beta_{5}\)) are constant over time.
